---
title: "Homework 3"
author: "JR Chansakul"
date: 2020-10-05
output: github_document
  
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)`columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 


How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
  arrange(desc(n))
```


Let's make a plot.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
  mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
    ) %>% 
  	ggplot(aes(x = aisle, y = n)) + 
    geom_point() +
  	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!


```{r}
instacart %>% 
    filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)

  
```

## Problem 2

## Problem 2 part 1 

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r Load_Tidy_Wrangle_Accel_dataset}

##Tidy and Wrangle Dataset
tidy_accel_df=
  read.csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
          activity_1:activity_1440,
          names_to = "activity_minute", 
          names_prefix = "activity_",
          values_to = "activity_count") %>% 
  mutate(
    Weekdays = 
      case_when(day %in% c("Saturday", "Sunday") ~ "weekend",
                day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~  "weekday")) %>%
  mutate(
      activity_minute = as.numeric(activity_minute),
      day = factor(day),
      day = fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  view ()

# View variable types of final dataset

view(tidy_accel_df) %>% view ()
sapply(tidy_accel_df, class) %>% view ()
```

The accelerometer dataset contains daily information of “activity counts” in one-minute intervals of a 63 year-old male with BMI 25, who was admitted to Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The tidied dataset contains `r ncol(tidy_accel_df)` variables and `r nrow(tidy_accel_df)` rows or observations.The variables of included in the dataset are `r names(tidy_accel_df)` and their variable types are `r sapply(tidy_accel_df, class)`, respectively.

We have `r n_distinct(pull(tidy_accel_df, day_id))` days of activity count data collected from the 65 year-old male by minute. We have the `week` variable that denotes the week that that accelerometer data was collected and the newly generated factor variable of `weekdays`, which denotes whether the data collected was on a weekend or a weekday.

## Problem 2 part 2

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r Daily_Activity_Count}

Total_activity_day = 
  tidy_accel_df %>% 
    group_by(week, day) %>%
  mutate(
    sum_activity_day = sum(activity_count),
    sum_activity_day = as.numeric(sum_activity_day)) %>% 
  select(week, day, sum_activity_day) %>% 
  distinct() %>%
  arrange (week, day) %>%
  relocate (week, day) %>%
  view ()

# View in table format 
Total_activity_day %>% 
knitr::kable()

# View variable types in dataset 
sapply(Total_activity_day, class) %>% view ()
```

The 63 year old male has a mean activity count per day of `r mean(pull(Total_activity_day, sum_activity_day))` with a standard deviation of `r sd(pull(Total_activity_day, sum_activity_day))` that ranges from `r min(pull(Total_activity_day, sum_activity_day))` to `r max(pull(Total_activity_day, sum_activity_day))`. His activity count seems to increase the first two weeks and then decline for the next two weeks. On week 5, the activity count seems to increase again. However, the two Saturdays on weeks 4 and 5 were his lowest activity count days at 1440. 

## Problem 2 part 3

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r Activity_Per_Day_Plot}
tidy_accel_df %>% 
  ggplot(aes(x = activity_minute, y = activity_count, color = day)) + 
  geom_line() + 
  labs(
        title = "Activity Level Per Minute",
        x = "Time",
        y = "Level of Activity") +
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("12AM","6AM","12PM", "6PM","12AM"),
    limits = c(0, 1440)) +
    viridis::scale_color_viridis(
      name = "Day",
      discrete = TRUE)+ 
  theme_set(theme_minimal() + theme(legend.position = "bottom")) +   
  theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1)) + 
  facet_grid(. ~ day)

ggsave("./plots/accel_plot.pdf", height = 4, width = 6) 
```

We can conclude from this graphic that our patient had increased activity between 6:00pm and 10:00pm on Monday, Wednesday, Friday and Saturday suggesting he may go on exercise on socializing with friends on these evenings. The patient seems to have morning activites between 6:00am and 9:00am from Wednesday, Thursday and Saturday. Additionally, the lowest acitvity times are between 11:00pm and 6:00am suggesting that the patient gets roughly 6-7 hours of sleep or downtime a week. 

## Problem 3
```{r}
noaa_data <- p8105.datasets::ny_noaa %>% 
  janitor::clean_names() 
```






y format to long format 
group by and summarize problem
geomline to connect the dots 
x minutes on x-axissand activity count on y-axis 
only 1 plot 





data manipulation steps and then plotting 


