---
title: "Homework 3"
author: "JR Chansakul"
date: 2020-10-05
output: github_document
  
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)`columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 


How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
  arrange(desc(n))
```


Let's make a plot.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
  mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
    ) %>% 
  	ggplot(aes(x = aisle, y = n)) + 
    geom_point() +
  	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!


```{r}
instacart %>% 
    filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)

  
```

## Problem 2

## Problem 2 part 1 

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r Load_Tidy_Wrangle_Accel_dataset}

##Tidy and Wrangle Dataset
tidy_accel_df=
  read.csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
          activity_1:activity_1440,
          names_to = "activity_minute", 
          names_prefix = "activity_",
          values_to = "activity_count") %>% 
  mutate(
    Weekdays = 
      case_when(day %in% c("Saturday", "Sunday") ~ "weekend",
                day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~  "weekday")) %>%
  mutate(
      activity_minute = as.numeric(activity_minute),
      day = factor(day),
      day = fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  view ()

# View variable types of final dataset

view(tidy_accel_df) %>% view ()
sapply(tidy_accel_df, class) %>% view ()
```

The accelerometer dataset contains daily information of “activity counts” in one-minute intervals of a 63 year-old male with BMI 25, who was admitted to Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The tidied dataset contains `r ncol(tidy_accel_df)` variables and `r nrow(tidy_accel_df)` rows or observations.The variables included in the dataset are `r names(tidy_accel_df)` and their variable types are `r sapply(tidy_accel_df, class)`, respectively.

We have `r n_distinct(pull(tidy_accel_df, day_id))` days of activity count data collected from the 65 year-old male by minute. We have the `week` variable that denotes the week that that accelerometer data was collected and the newly generated factor variable of `weekdays`, which denotes whether the data collected was on a weekend or a weekday.

## Problem 2 part 2

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r Daily_Activity_Count}

Total_activity_day = 
  tidy_accel_df %>% 
    group_by(week, day) %>%
  mutate(
    sum_activity_day = sum(activity_count),
    sum_activity_day = as.numeric(sum_activity_day)) %>% 
  select(week, day, sum_activity_day) %>% 
  distinct() %>%
  arrange (week, day) %>%
  relocate (week, day) %>%
  view ()

# View in table format 
Total_activity_day %>% 
knitr::kable()

# View variable types in dataset 
sapply(Total_activity_day, class) %>% view ()
```

The 63 year old male has a mean activity count per day of `r mean(pull(Total_activity_day, sum_activity_day))` with a standard deviation of `r sd(pull(Total_activity_day, sum_activity_day))` that ranges from `r min(pull(Total_activity_day, sum_activity_day))` to `r max(pull(Total_activity_day, sum_activity_day))`. His activity count seems to increase the first two weeks and then decline for the next two weeks. On week 5, the activity count seems to increase again. However, the two Saturdays on weeks 4 and 5 were his lowest activity count days at 1440. 

## Problem 2 part 3

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r Activity_Per_Day_Plot}
tidy_accel_df %>% 
  ggplot(aes(x = activity_minute, y = activity_count, color = day)) + 
  geom_line() + 
  labs(
        title = "Activity Level Per Minute",
        x = "Time",
        y = "Level of Activity") +
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("12AM","6AM","12PM", "6PM","12AM"),
    limits = c(0, 1440)) +
    viridis::scale_color_viridis(
      name = "Day",
      discrete = TRUE)+ 
  theme_set(theme_minimal() + theme(legend.position = "bottom")) +   
  theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1)) 

ggsave("./plots/accel_plot.pdf", height = 4, width = 6) 
```

We can conclude from this busy graphic that our patient had increased activity between 6:00pm and 10:00pm on Monday, Wednesday, Friday and Saturday suggesting he may go on exercise on socializing with friends on these evenings. The patient seems to have morning activites between 6:00am and 9:00am from Wednesday, Thursday and Saturday. Additionally, the lowest acitvity times are between 11:00pm and 6:00am suggesting that the patient gets roughly 6-7 hours of sleep or downtime a week. 

## Problem 3

## Problem 3 Part A

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. 

```{r Import&Summarize_NOAA}

# Import
data("ny_noaa")

# Looking at distinct weather station IDs and number of observations
ny_noaa %>% 
  group_by(id) %>% 
  summarize (n())

# Date Range 
date_summary = ny_noaa %>% 
  group_by(date) %>% 
  summarize(n())  
head(date_summary, 5)
tail(date_summary, 5)

# Percentage of NA's by column
colMeans(is.na(ny_noaa))  

# Varialbe Types in dataset
sapply(ny_noaa, class) %>% view ()
```

The original NOAA National Climatic Data Center dataset has `r ncol(ny_noaa)` variables and `r nrow(ny_noaa)` observations/rows. The variables included in the dataset are `r names(ny_noaa)` and their variable type is `r sapply(ny_noaa, class)`, respectively.

Below is a brief description of the original variables:

id: Weather station ID
date: Date of observation
prcp: Precipitation (tenths of mm)
snow: Snowfall (mm)
snwd: Snow depth (mm)
tmax: Maximum temperature (tenths of degrees C)
tmin: Minimum temperature (tenths of degrees C)

There are 747 unique weather stations that may provide information on precipitation, snowfall, snow depth, min and max temperature on any given day from 1981-01-01 to 2010-21-31. Unfortunately, we have a lot of missing data. In particular, we found that the `tmin` and `tmax` columns have 43.71% of its observations as NA and 22.80% of the observations for the `snwd` variable as NA. 

## Problem 3 Part B

Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r Clean_NY_NOAA}

#Clean Dataset and Units
noaa_data = ny_noaa %>% 
  janitor::clean_names() %>% 
    separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    prcp = prcp/10,
    snow = as.numeric(snow),
    snwd = as.numeric(snwd),
    tmax = as.numeric(tmax)/10,
    tmin = as.numeric(tmin)/10)

# Commonly observed Snowfall values
noaa_data %>% 
  count(snow) %>% 
  arrange(desc(n))
```

I separated date into year, month and day as well as converted precipitation to mm, tmax to C, and tmin to C by dividing by 10. I found that the most commonly observed snowfall  value is 0. This is most likely due to snow only occurring in the winter. 


## Problem 3 Part C

Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?


```{r}
noaa_data %>% 
  
  ggplot(aes(x = year, y = tmax, fill = month)) + 
  geom_line() + 
  labs(
        title = "Average Max Temp in each station across years",
        x = "Year",
        y = "Average Max Temperature (C)")
```




## Problem 3 Part D

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}

```


x minutes on x-axissand activity count on y-axis 
only 1 plot 
data manipulation steps and then plotting 


