---
title: "Homework 3"
author: "JR Chansakul"
date: 2020-10-05
output: github_document
  
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)`columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 


How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
  arrange(desc(n))
```


Let's make a plot.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
  mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
    ) %>% 
  	ggplot(aes(x = aisle, y = n)) + 
    geom_point() +
  	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!


```{r}
instacart %>% 
    filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)

  
```

## Problem 2

## Problem 2 part 1 

```{r Load_Tidy_Wrangle_Accel_dataset}

##Tidy and Wrangle Dataset
tidy_accel_df=
  read.csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
          activity_1:activity_1440,
          names_to = "activity_minutes", 
          names_prefix = "activity_",
          values_to = "activity_count") %>% 
  mutate(Weekdays = 
           case_when(day %in% c("Saturday", "Sunday") ~ "weekend",
                     day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~  "weekday"),
           day = fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  mutate(activity_minutes = as.numeric(activity_minutes)) %>%
  view ()

# View variable types of final dataset

view(tidy_accel_df)
sapply(tidy_accel_df, class) 
```






y format to long format 
group by and summarize problem
geomline to connect the dots 
x minutes on x-axissand activity count on y-axis 
only 1 plot 



## Problem 3

data manipulation steps and then plotting 


